---
title: "The day the carnival came to town"
author: "David Hood, @thoughtfulnz, github.com/thoughtfulbloke/kiwicon2038"
date: "12/11/2018"
output:
  powerpoint_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(dpi = 280)
```

```{r libraries, message=FALSE, warning=FALSE}
library(readr)
library(dplyr)
library(lubridate)
library(ggplot2)
library(purrr)
library(tidyr)
library(ggraph)
library(igraph)
library(ggthemes)
library(gganimate)
library(patchwork)
library(tidytext)
library(stringr)
library(ggrepel)
```

```{r coredata}
# coredata has v_for visitor (tweet discussion) and
# m_for_mentions (notifications from the tweets)
load("~/Syncplicity Folders/support_files/coredata.RData")
n_for_nznetwork <- read.csv("~/Syncplicity Folders/support_files/nzmatches.csv",
                          colClasses = "character") # nz friend relationships
r_for_radicalism <- read.csv("~/Syncplicity Folders/support_files/fa_spec.csv",
                          colClasses = "character") # were on spectrum account
c_for_country <- read.csv("~/Syncplicity Folders/support_files/assignment.csv",
                          colClasses = "character")
f_for_friends <- read.csv("~/Syncplicity Folders/support_files/friends.csv",
                          colClasses = "character") %>%
    mutate(lookup = paste(paste0("x",user), paste0("x",friends_id)))
```

## Dataset

- From targetted searches & capturing at time
- 2018-07-25 13:03:40 to 2018-08-06 23:45:21 UTC
- 115218 tweets, 3065709 mentions
- 373693 accounts loosely involved
- 40798 accounts tightly involved
- 53900649 friends relationships
- 21366303 likes

## Backstory

```{r, fig.height=3, dpi=300}
# 5th July
data_as_txt <- "
state, opposed, supportive
NZ, 25.5, 81.3
AU, 25.5, 0
US, 20.7, 12.5
UK, 10.9, 6.3
robots, 7.6, 0
anon, 8.2, 0
other, 1.6, 0"

read.csv(text=data_as_txt) %>% gather(status, percent, 2:3) %>%
    ggplot(aes(x=status, y=percent, fill=state)) + 
    geom_col(width=0.3) + coord_flip() + theme_minimal() +
    scale_fill_colorblind() + xlab("reaction to Mayor")

```

## Composition

```{r}
c_for_country %>% 
    mutate(location = case_when(
        locat == "offshore" ~ "not nz",
        locat == "US" ~ "US",
        locat == "" ~ "unknown",
        locat == "GB" ~ "UK",
        locat == "AU" ~ "AU",
        locat == "NZ" ~ "NZ",
        locat == "CA" ~ "CA",
        TRUE ~ "other"
    )) %>%
    count(location, sort = TRUE) %>%
    ggplot(aes(x=location, y=n, fill=location)) + 
    geom_col(width=0.3) + theme_minimal() +
    scale_fill_colorblind() + xlab("inferred location")
```

## Politics - notifiers

```{r, fig.height=5}
c_for_country <- c_for_country %>% mutate(xid=paste0("x", user_id))
grf <- v_for_visitor %>%
    count(user_id) %>%
    inner_join(r_for_radicalism, by=c("user_id" = "xliker")) %>%
    inner_join(c_for_country, by=c("user_id" = "xid")) %>% 
    rename(place = locat) %>%
    filter(place %in% c("NZ","AU","GB","US")) %>%
    mutate(
        balance = as.numeric(nett_fa) - 
            (as.numeric(nett_nz) + as.numeric(nett_au) + 
                 as.numeric(nett_gb) + as.numeric(nett_us)),
        focus = (as.numeric(nett_fa) + as.numeric(nett_nz) + as.numeric(nett_au) + 
                 as.numeric(nett_gb) + as.numeric(nett_us))/as.numeric(allike)
    ) %>%
    count(user_id, place, balance,focus, n) 
    ggplot(grf, aes(x = balance, y=log2(focus + 0.01), colour=place)) +
        geom_vline(xintercept = 0, alpha=0.4) + geom_jitter(alpha=0.3) +
     facet_wrap(~ place, nrow=2, strip.position="bottom") + scale_colour_colorblind() + theme_void() +
        theme(legend.position="none", strip.text.x = element_text(size = 32))
        
    

```


## Community


```{r}

m_for_mentions %>%
    mutate(nox=gsub("x", "", mentions)) %>%
    inner_join(c_for_country, by=c("nox"="user_id")) %>%
    filter(locat == "NZ") %>%
    mutate(close = !paste(mentions, user_id) %in% f_for_friends$lookup,
           NZ_time = with_tz(ymd_hms(created_at), tz="Pacific/Auckland"),
           step1 = floor_date(NZ_time,unit = "hour")) %>%
    group_by(mentions, step1) %>% 
    summarise(og = sum(!close), ogthres = og > 10) %>%
    group_by(step1) %>% summarise(thres=sum(ogthres)) %>%
    ggplot(aes(x=step1, y=thres)) + geom_line() + theme_minimal() +
    ggtitle("Of 1505 NZ accounts, the number of accounts in a given hour \ngettting more than 10 notifications from outside of thier friend groups") +
    xlab("Date") + ylab("Number of accounts") +
    geom_vline(xintercept = ISOdatetime(2018,8,2,19,44,00, tz="Pacific/Auckland"), 
               colour="blue", lwd=1.1) +
    annotate("text", x = ISOdatetime(2018,8,3,9,44,00, tz="Pacific/Auckland"),
             y = 30, label="visit", colour="blue")
```

## Time

```{r}
v_for_visitor %>% inner_join(c_for_country, by=c("user_id" = "xid")) %>%
    filter(locat %in% c("AU", "NZ", "US", "GB")) %>%
    mutate(hour_of_day = hour(created_at)) %>% 
    count(locat, hour_of_day) %>% 
    group_by(locat) %>%
    mutate(prop = n/sum(n)) %>%
    ggplot(aes(x=hour_of_day, y=prop, fill=locat)) + 
    facet_wrap(~locat, ncol=2) +
    scale_colour_colorblind() +
    geom_col() + theme_void() + coord_polar() +
    theme(legend.position="none", strip.text.x = element_text(size = 28)) 
```

## Language (1) Rate difference

```{r}
nz <- v_for_visitor %>% 
    left_join(c_for_country, by=c("user_id" = "xid")) %>%
    filter(locat == "NZ") %>% select(status_id, text, locat)
us <- v_for_visitor %>% 
    left_join(c_for_country, by=c("user_id" = "xid")) %>%
    filter(locat == "US") %>% select(status_id, text, locat)
bind_rows(nz,us) %>%
    unnest_tokens(word, text, token = "tweets") %>%
    count(locat,word) %>%
    filter(!str_detect(word,fixed("@"))) %>% 
    filter(!str_detect(word,fixed("#"))) %>%
    group_by(locat) %>%
    mutate(prop = n/sum(n)) %>%
    ungroup() %>%
    filter(prop > 0.0001) %>%
    select(-n) %>%
    spread(locat, prop,fill = 0) %>%
    filter(NZ != 0, US != 0) %>%
    mutate(ratio = US/NZ) %>% 
    arrange(ratio) %>%
    slice(c(1:13,(n()-12):n())) %>% 
    mutate(xval = ifelse(ratio < 1, -1 * NZ/US, ratio),
           yval = ifelse(ratio < 1, NZ, US),
           c_is = ifelse(ratio < 1, "NZ", "US")) %>%
    ggplot(aes(x=xval, y=yval, label=word, colour=c_is)) + 
    geom_text_repel(size=6) +
    geom_point() + geom_vline(xintercept = 0, alpha=0.3) + theme_void() +
    ylim(-0.0005, NA) +
    theme(legend.position="none", strip.text.x = element_text(size = 28)) +
    scale_colour_colorblind()
```

## Language (2) Absolute difference

```{r}
nz <- v_for_visitor %>% 
    left_join(c_for_country, by=c("user_id" = "xid")) %>%
    filter(locat == "NZ") %>% select(status_id, text, locat)
us <- v_for_visitor %>% 
    left_join(c_for_country, by=c("user_id" = "xid")) %>%
    filter(locat == "US") %>% select(status_id, text, locat)
abswords <- bind_rows(nz,us) %>%
    unnest_tokens(word, text, token = "tweets") %>%
    anti_join(get_stopwords(), by="word") %>%
    count(locat,word) %>%
    filter(!str_detect(word,fixed("@"))) %>% 
    filter(!str_detect(word,fixed("#"))) %>%
    group_by(locat) %>%
    mutate(prop = n/sum(n)) %>%
    ungroup() %>%
    filter(prop > 0.0001) %>%
    select(-n) %>%
    spread(locat, prop,fill = 0) %>%
    mutate(ratio = US-NZ) %>% 
    arrange(ratio) %>%
    slice(c(1:13,(n()-12):n())) %>% 
    mutate(xval = ifelse(ratio < 0, -1 * NZ, ratio),
           yval = ifelse(ratio < 0, NZ, US),
           c_is = ifelse(ratio < 0, "NZ", "US"))
abswords$word[18] <- "USFLAG"
ggplot(abswords, aes(x=xval, y=yval, label=word, colour=c_is)) + 
    geom_text_repel(size=6) +
    geom_point() + geom_vline(xintercept = 0, alpha=0.3) + theme_void() +
    theme(legend.position="none", strip.text.x = element_text(size = 28)) +
    scale_colour_colorblind()
```

## Uneven

```{r}

m_for_mentions %>% 
    mutate(NZ_time = with_tz(ymd_hms(created_at), tz="Pacific/Auckland"),
           step1 = floor_date(NZ_time,unit = "hour")) %>%
    filter(step1 >= ISOdatetime(2018,8,1,0,0,0, tz="Pacific/Auckland"),
           step1 < ISOdatetime(2018,8,4,0,0,0, tz="Pacific/Auckland")) %>%
    inner_join(c_for_country, by=c("user_id" = "xid")) %>% 
    rename(uloc = locat) %>%
    inner_join(c_for_country, by=c("mentions" = "xid")) %>% 
    filter(locat == "NZ") %>%
    group_by(mentions) %>%
    summarise(n_mentions = n(), 
              percent_NZ = 100 *sum(uloc == "NZ")/n()) %>%
    ggplot(aes(x=n_mentions, y=percent_NZ)) + 
    geom_jitter(alpha=0.3) +
    theme_minimal() + xlab("number of mentions") +
    ylab("percent of NZ origin")

```

## Possible Responses

- all along the watchtower
- pick on someone your own size
- burn it all down